---
title: "Raman analysis; LGraft Score prediction"
author: "Maëlle"
date: "`r Sys.Date()`"
output:
  html_document:
    fig_width: 15
    fig_height: 10
    toc: yes
    theme: simplex
    keep_md: yes
    always_allow_html: true
  pdf_document:
    toc: yes
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
#rm(list = ls()) 
source("Z:/PROJETS RAMAN/Code/Code greffon hépatique/Raman_transplant/functions_raman.R")
knitr::opts_chunk$set(echo = F, warning = FALSE, message = FALSE)
```

```{r}
library(rstatix)
library(ggpubr)
```

## Data preprocessing

```{r Data importation}
data <- import_sp('TOUS/', 1157)

### Supplementary information ###
info <- read_excel("Z:/PROJETS RAMAN/Code/Code greffon hépatique/Raman_transplant/cleaned_data_R.xlsx")
info$l7_high_low <- ifelse(info$L7_groupe %in% c("Gpe1", "Gpe2", "Gpe3"), "Low-risk", "High-risk")
info$EAD <- ifelse(info$EAD == 1, "EAD", "no EAD")
```

Some spectra are too noisy to be considered as accurate. The calculation of the 2nd Savitsky-Golay derivative allows the identification of spectra with many oscillations but no real peaks.To quantify the oscillations: the mean and standard-deviation of the 2nd Savitsky-Golay derivative were calculated. Spectra with an high mean and a low standard-deviation were removed from the dataset.

```{r}
data2 <- filter_noised_spectra(data)
```

Only two regions of interest in the spectra were selected:

-   Z1: The 700-1800 region, corresponding to the sample fingerprint.

-   Z2: The 2800-300 region, corresponding to the C-H groups, also named high wavenumber region

-   The remaining zones are called silent zones, the signal contains mostly noise.

For all the spectra, these two regions were normalized using the standard normal variate method.

```{r}
### SPLIT BY ZONE ###
z1 <- data2[,220:700]
z1_n <- normalize(z1)

z2 <- data2[,1300:1450]
z2_n <- normalize(z2)
```

## Correlation with L-GrAFT7 score

```{r}
groups <- get_varY(data2, info, 'L7_groupe')

avg1 <- compute_average_factor(z1_n, groups)
plot_spectra1(avg1, titre="Average Raman spectra (fingerprint region) across L-GrAFT7 score groups")


plot_spectra1(avg1)

avg2 <- compute_average_factor(z2_n, groups)
plot_spectra1(avg2, titre="Average Raman spectra (high wavenumber region) across L-GrAFT7 score groups")
plot_spectra1(avg2)
```

```{r}
ratio_lipids_prot <- ratio_2_pics(z2_n, 2928, 2854)
ratio_lipids_prot$Group <- groups

ggplot(ratio_lipids_prot, aes(x=Group, y=`Ratio 2928 on 2854`, fill=Group)) + 
  geom_boxplot(outlier.shape = NA) + scale_y_discrete()  + stat_compare_means(label.y.npc = "bottom", label.x.npc = "center") +   theme_classic() +   theme(     axis.title.x = element_text(size=14, face="bold", colour = "black"),         axis.title.y = element_text(size=14, face="bold", colour = "black"),         axis.text.x = element_text(size=12, face="bold", colour = "black"),      axis.text.y = element_text(size=12, face="bold", colour = "black"),     strip.text.x = element_text(size = 10, face="bold", colour = "black" ),     strip.text.y = element_text(size = 10, face="bold", colour = "black")   ) + ggtitle("Distribution of Proteins/Lipids ratio values across L-GrAFT7 score groups")

ggplot(ratio_lipids_prot, aes(x=Group, y=`Ratio 2928 on 2854`, fill=Group)) + 
  geom_boxplot(outlier.shape = NA) + scale_y_discrete() + stat_compare_means(label.y.npc = "bottom", label.x.npc = "center") +
  theme_classic() +
  theme(
    axis.title.x = element_text(size=14, face="bold", colour = "black"),    
    axis.title.y = element_text(size=14, face="bold", colour = "black"),    
    axis.text.x = element_text(size=12, face="bold", colour = "black"), 
    axis.text.y = element_text(size=12, face="bold", colour = "black"),
    strip.text.x = element_text(size = 10, face="bold", colour = "black" ),
    strip.text.y = element_text(size = 10, face="bold", colour = "black")
  )

```




## Correlation with L-GrAFT10 score

```{r}
groups <- get_varY(data2, info, 'L10_gpe')

avg1 <- compute_average_factor(z1_n, groups)
plot_spectra1(avg1, titre="Average Raman spectra (fingerprint region) across L-GrAFT10 score groups")


plot_spectra1(avg1)

avg2 <- compute_average_factor(z2_n, groups)
plot_spectra1(avg2, titre="Average Raman spectra (high wavenumber region) across L-GrAFT10 score groups")

plot_spectra1(avg2)
```

```{r}
ratio_lipids_prot <- ratio_2_pics(z2_n, 2928, 2854)
ratio_lipids_prot$Group <- groups

ggplot(ratio_lipids_prot, aes(x=Group, y=`Ratio 2928 on 2854`, fill=Group)) + 
  geom_boxplot(outlier.shape = NA) + scale_y_discrete() + ggtitle("Distribution of Proteins/Lipids ratio values across L-GrAFT10 score groups")

ggplot(ratio_lipids_prot, aes(x=Group, y=`Ratio 2928 on 2854`, fill=Group)) + 
  geom_boxplot(outlier.shape = NA) + scale_y_discrete()  + stat_compare_means(label.y.npc = "bottom", label.x.npc = "center") +   theme_classic() +   theme(     axis.title.x = element_text(size=14, face="bold", colour = "black"),         axis.title.y = element_text(size=14, face="bold", colour = "black"),         axis.text.x = element_text(size=12, face="bold", colour = "black"),      axis.text.y = element_text(size=12, face="bold", colour = "black"),     strip.text.x = element_text(size = 10, face="bold", colour = "black" ),     strip.text.y = element_text(size = 10, face="bold", colour = "black")   ) 


```

## Correlation with High risk - low Risk

High risk: L7 Score groups : 4, 5, 6, 7
Low risk: L7 Score groups : 1, 2, 3

```{r}
groups <- get_varY(data2, info, 'l7_high_low')

avg1 <- compute_average_factor(z1_n, groups)
plot_spectra1(avg1, c("royalblue", "tomato"), titre="Average Raman spectra (fingerprint region) across l7_high_low groups")

plot_spectra1(avg1, c("royalblue", "tomato"))

avg2 <- compute_average_factor(z2_n, groups)
plot_spectra1(avg2, c("royalblue", "tomato"), titre="Average Raman spectra (high wavenumber region) across l7_high_low groups")

plot_spectra1(avg2, c("royalblue", "tomato"))
```

```{r}
ratio_lipids_prot <- ratio_2_pics(z2_n, 2928, 2854)
ratio_lipids_prot$Group <- groups

ggplot(ratio_lipids_prot, aes(x=Group, y=`Ratio 2928 on 2854`, fill=Group)) + 
  geom_boxplot(outlier.shape = NA) + scale_y_discrete()  + stat_compare_means(label.y.npc = "bottom", label.x.npc = "center") +   theme_classic() +   theme(     axis.title.x = element_text(size=14, face="bold", colour = "black"),         axis.title.y = element_text(size=14, face="bold", colour = "black"),         axis.text.x = element_text(size=12, face="bold", colour = "black"),      axis.text.y = element_text(size=12, face="bold", colour = "black"),     strip.text.x = element_text(size = 10, face="bold", colour = "black" ),     strip.text.y = element_text(size = 10, face="bold", colour = "black")   ) + ggtitle("Distribution of Proteins/Lipids ratio values across l7_high_low groups")  + scale_fill_manual(values=c("tomato", "royalblue"))

ggplot(ratio_lipids_prot, aes(x=Group, y=`Ratio 2928 on 2854`, fill=Group)) + 
  geom_boxplot(outlier.shape = NA) + scale_y_discrete()  + stat_compare_means(label.y.npc = "bottom", label.x.npc = "center") +   theme_classic() +   theme(     axis.title.x = element_text(size=14, face="bold", colour = "black"),         axis.title.y = element_text(size=14, face="bold", colour = "black"),         axis.text.x = element_text(size=12, face="bold", colour = "black"),      axis.text.y = element_text(size=12, face="bold", colour = "black"),     strip.text.x = element_text(size = 10, face="bold", colour = "black" ),     strip.text.y = element_text(size = 10, face="bold", colour = "black")   ) +  scale_fill_manual(values=c("tomato", "royalblue"))

```

## Correlation with EAD/ no EAD



```{r}
groups <- get_varY(data2, info, 'EAD')
# groups <- factor(groups, levels = c('EAD', 'no EAD'))

avg1 <- compute_average_factor(z1_n, groups)
plot_spectra1(avg1, c("royalblue", "tomato"), titre="Average Raman spectra (fingerprint region) across EAD groups")

avg2 <- compute_average_factor(z2_n, groups)
plot_spectra1(avg2, c("royalblue", "tomato"), titre="Average Raman spectra (high wavenumber region) across EAD groups")
```

```{r}
ratio_lipids_prot <- ratio_2_pics(z2_n, 2928, 2854)
ratio_lipids_prot$Group <- groups

ggplot(ratio_lipids_prot, aes(x=Group, y=`Ratio 2928 on 2854`, fill=Group)) + 
  geom_boxplot(outlier.shape = NA) + scale_y_discrete()  + stat_compare_means(label.y.npc = "bottom", label.x.npc = "center") +   theme_classic() +   theme(     axis.title.x = element_text(size=14, face="bold", colour = "black"),         axis.title.y = element_text(size=14, face="bold", colour = "black"),         axis.text.x = element_text(size=12, face="bold", colour = "black"),      axis.text.y = element_text(size=12, face="bold", colour = "black"),     strip.text.x = element_text(size = 10, face="bold", colour = "black" ),     strip.text.y = element_text(size = 10, face="bold", colour = "black")   )  + scale_fill_manual(values=c("tomato", "royalblue")) + ggtitle("Distribution of Proteins/Lipids ratio values across EAD groups")

ggplot(ratio_lipids_prot, aes(x=Group, y=`Ratio 2928 on 2854`, fill=Group)) + 
  geom_boxplot(outlier.shape = NA) + scale_y_discrete()  + stat_compare_means(label.y.npc = "bottom", label.x.npc = "center") +   theme_classic() +   theme(     axis.title.x = element_text(size=14, face="bold", colour = "black"),         axis.title.y = element_text(size=14, face="bold", colour = "black"),         axis.text.x = element_text(size=12, face="bold", colour = "black"),      axis.text.y = element_text(size=12, face="bold", colour = "black"),     strip.text.x = element_text(size = 10, face="bold", colour = "black" ),     strip.text.y = element_text(size = 10, face="bold", colour = "black")   )  + scale_fill_manual(values=c("tomato", "royalblue"))

```

